---
title: "Wrangling Competition Data"
author: "Katherine Burley & Maggie O'Shea"
output: pdf_document
editor_options: 
  chunk_output_type: console
---



## Load Libraries

```{r message = FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(kableExtra)
```

## Import the Dataset

#### Instructions:

In the folder `/Competition/Data` you will find three datasets one with hourly demand, one with hourly temperature and another with relative humidity from January 2005 to December 2010.
Your goal is to forecast **daily** demand for the month of January 2011 based on this historical data. You may or may not use the temperature and relative humidity in your models. The temperature and humidity measurement are from stations close to the household meter data you have.


```{r}
hourly_demand <- read_excel("./Data/load.xlsx")
hourly_temp<- read_excel("./Data/temperature.xlsx")
hourly_humidity<- read_excel("./Data/relative_humidity.xlsx")
submission_template <- read_excel("./Data/submission_template.xlsx")
```


## Wrangle/Process the Dataset

#### Instructions: 

You will need to transform hourly data into daily data. See the Rmd file from Lesson 11 for instructions on how to aggregate your dataset using pipes.

Note that I provided hourly data. You should take the **average** of the 24 hours to obtain the daily load.


```{r}
# daily_demand <- hourly_demand%>%
#   group_by(date, meter_id)%>%
#   summarize(
#     daily_average = (h1+h2+h3+h4+h5+h6+h7+h8+h9+h10+h11+h12+
#                        h13+h14+h15+h16+h17+h18+h19+h20+h21+h22+h23+h24)/24
#   )

daily_demand <- hourly_demand %>%
  pivot_longer(cols=h1:h24, names_to = "hour", values_to = "demand") %>%
  filter(!is.na(demand)) %>% # only a handful of missing hourly obs. Let's try removing them
  group_by(date, meter_id) %>%
  summarise(daily_average = mean(demand))

daily_temp <- hourly_temp%>%
  group_by(date)%>%
  summarize(avg_t_ws1 = mean(t_ws1),
            avg_t_ws2 = mean(t_ws2),
            avg_t_ws3 = mean(t_ws3),
            avg_t_ws4 = mean(t_ws4),
            avg_t_ws5 = mean(t_ws5),
            avg_t_ws6 = mean(t_ws6),
            avg_t_ws7 = mean(t_ws7),
            avg_t_ws8 = mean(t_ws8),
            avg_t_ws9 = mean(t_ws9),
            avg_t_ws10 = mean(t_ws10),
            avg_t_ws11 = mean(t_ws11),
            avg_t_ws12 = mean(t_ws12),
            avg_t_ws13 = mean(t_ws13),
            avg_t_ws13 = mean(t_ws13),
            avg_t_ws14 = mean(t_ws14),
            avg_t_ws15 = mean(t_ws15),
            avg_t_ws16 = mean(t_ws16),
            avg_t_ws17 = mean(t_ws17),
            avg_t_ws18 = mean(t_ws18),
            avg_t_ws19 = mean(t_ws19),
            avg_t_ws20 = mean(t_ws20),
            avg_t_ws21 = mean(t_ws21),
            avg_t_ws22 = mean(t_ws22),
            avg_t_ws23 = mean(t_ws23),
            avg_t_ws24 = mean(t_ws24),
            avg_t_ws25 = mean(t_ws25),
            avg_t_ws26 = mean(t_ws26),
            avg_t_ws27 = mean(t_ws27),
            avg_t_ws28 = mean(t_ws28))

daily_humidity <- hourly_humidity%>%
  pivot_longer(cols = -c(hr, date), names_to = "rh", values_to = "humidity")%>%
  group_by(date, rh)%>%
  summarize(avg_humidity = mean(humidity))%>%
  pivot_wider(values_from = "avg_humidity", names_from = "rh")

# Extract forecast dates from the submission template
forecast_dates <- submission_template$date
rm(cor_mat, daily_all, daily_forcor, daily_forcor_mat, submission_template)
  
```


## Create a Time Series Object

#### Instructions: 

After you process your dataset use the `msts()` function to create a time series object. You need to use `msts()` instead of `ts()` because your daily data will have more than one seasonal component.

```{r}
# Look at data
ggplot(daily_demand) +
  geom_line(aes(x=daily_demand$date, y=daily_demand$daily_average))

# Create ts object
ts_daily_demand <- msts(daily_demand$daily_average, 
                           seasonal.periods =c(7, 365.25),
                           start=c(2005,01,01))

ts_daily_humidity <- msts(daily_humidity$rh_ws1, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))
ts_daily_temp <- msts(daily_temp$avg_t_ws1, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

# **NOTE**: Would be nice to add monthly, but I don't see this used in the M10 code or the link referenced there. I'm guessing probably because months can't be expressed precisely in # of days
## I made monthly data! Though because there are so few years, when you reduce to monthly data there are so few observations which makes me think it is better to model with daily data? Open to your thoughts though!
monthly_demand <- daily_demand%>%
  mutate("month" = month(ymd(date)),
         "year" = year(ymd(date)))%>%
  select(month, year, daily_average)%>%
  group_by(month, year)%>%
  summarize(monthly_average = mean(daily_average))%>%
  mutate("month-year" = paste0(month, year, sep="/"))

ggplot(monthly_demand) +
  geom_line(aes(x=`month-year`, y=monthly_average), group=1)
#There seems to be missing data which is a bit weird, could be because of the way I summarized it?
# updated the code in Chunk 3 to just remove the missing hourly demand obs? There are only like 5, and since its a daily average I think its ok to just drop?

# Create training and testing data from the full ts object
ts_demand_train <- subset(ts_daily_demand,
                          end = length(ts_daily_demand)-365)
ts_humid_train <- subset(ts_daily_humidity,
                          end = length(ts_daily_demand)-365)
ts_temp_train <- subset(ts_daily_temp,
                          end = length(ts_daily_demand)-365)
#create a subset for testing purpose
ts_demand_test <- subset(ts_daily_demand,
                         start = length(ts_daily_demand)-365)
ts_demand_test <- subset(ts_demand_test,
                         end = length(ts_demand_test)-1) 
# **NOTE**: also something odd going on here with the dates where we are getting an observation for 1/1/2011

ts_temp_test <- subset(ts_daily_temp,
                          start = length(ts_daily_demand)-365)
ts_temp_test <- subset(ts_temp_test,
                         end = length(ts_temp_test)-1) 

ts_humid_test <- subset(ts_daily_humidity,
                          start = length(ts_daily_demand)-365)
ts_humid_test <- subset(ts_humid_test,
                         end = length(ts_humid_test)-1) 

# Decompose 
ts_daily_demand %>% mstl() %>%
  autoplot()

decompose <- data.frame(mstl(ts_daily_demand))
ts_remainder <- ts(decompose$Remainder)

## Look for correlations between humidity, temp, and demand/de-seasoned demand
deseasoned <- data.frame(date = daily_demand$date, deseas_demand = ts_remainder)
daily_all <- daily_demand %>%
  left_join(deseasoned, by="date") %>%
  left_join(daily_humidity, by="date") %>%
  left_join(daily_temp, by="date") %>%
  ungroup()

daily_forcor <- daily_all %>%
  select(-c(date, meter_id))

daily_forcor_mat <- data.matrix(daily_forcor)

cor_mat <- cor(daily_forcor, use="complete")
cor_results <- data.frame(corr = cor_mat[,1:2])
cor_results <- cor_results %>%
  mutate(corr.daily_average = abs(corr.daily_average),
         corr.deseas_demand = abs(corr.deseas_demand))
  # none of the humidity or temp vars have a very strong correlation with demand or deseasoned demand...

# Plot the time series over time, ACF and PACF
par(mfrow=c(1,3), mar=c(6, 4, 4, 2))
plot(ts_remainder, main="Remainder Time Series", ylab = "Demand")
Acf(ts_remainder,lag.max=40, main="ACF") 
Pacf(ts_remainder,lag.max=40, main="PACF")
  # Looks like an AR process?

```


# Fit Models to your Data

Fit models to your dataset considering the period Jan 1st 2005 to Dec 31st 2009. First, we will consider an ARIMA with Fourier terms to account for multiple seasonalities in our data. For this specification, we assume that there are two sine and cosine terms during the weekly season and six sine and cosine terms during the annual season, after testing different combinations of K.  

## ARIMA with Fourier Terms Model 
``` {r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
# First Model: 
ARIMA_Fourier_fit <- auto.arima(ts_demand_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_demand_train, 
                                          K=c(2,6)) # 2 weekly cycles, 6 annual cycles
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Fourier_forecast <- forecast(ARIMA_Fourier_fit,
                           xreg=fourier(ts_demand_train,
                                        K=c(2,6),
                                        h=365),
                           h=365
                           ) 

#Plot forecasting results
autoplot(ARIMA_Fourier_forecast) + ylab("Demand")

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(ARIMA_Fourier_forecast, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Demand")

# Plot model and observed data during 2010
plot(ARIMA_Fourier_forecast$mean,
     xlab = "Date",
     ylab = "Demand",
     ylim = c(1000, 10000))
lines(ts_demand_test,
      type = "l",
      col = 4,
      lwd = 2)
legend(2010, 10000, legend=c("Forecasting Results", "Last Year of Data (2010)"),
       col=c("black",4), lwd=3, cex=0.8)

ARIMA_four_scores <- accuracy(ARIMA_Fourier_forecast$mean, ts_demand_test)
checkresiduals(ARIMA_Fourier_forecast)

```

### ARIMA with Fourier - Training Data Interpretation
THE RMSE is 951.8972 and the MAPE is 23.426. 

### Modeling Full Data - ARIMA with Fourier
After using the training dataset, re-running with the full dataset to forecast into 2011:
```{r}
# First Model: 
ARIMA_Fourier_full <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_daily_demand, 
                                          K=c(2,6)) # 2 weekly cycles, 2 annual cycles
                             )
n_forecast = 31+28 # January and February - 59 days

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Fourier_FULLforecast <- forecast(ARIMA_Fourier_full,
                           xreg=fourier(ts_daily_demand,
                                        K=c(2,6),
                                        h=n_forecast),
                           h=n_forecast
                           #Set H to 60 because I think we are only forecasting till February. I changed both h values to 60 because when only changing 1 (just the second one not in the xreg parantheses) it still had 365 days of forecasting. 
                           ) 
autoplot(ARIMA_Fourier_FULLforecast) + ylab("Demand")

```

### Export Model Results
```{r}
ARIMA_fourier_export <- data.frame(date=forecast_dates, load=ARIMA_Fourier_FULLforecast$mean)
write.csv(ARIMA_fourier_export, "./Submissions/ARIMA_fourier_v2.csv", row.names = FALSE)
```

## ARIMA with Exogenous Variables

### Both Temperature and Humidity 
```{r warnings = FALSE, message=FALSE}
tempandhumidity <- auto.arima(ts_demand_train,
                  seasonal = FALSE, 
                  lambda = 0,
         xreg = cbind(fourier(ts_demand_train, 
                                          K=c(2,6)),
               humidity=ts_humid_train,
               temperature=ts_temp_train))



#Forecast 
temphumidity_forecast <- forecast(tempandhumidity,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),h=365),
                                humidity=ts_humid_test,
                                temperature=ts_temp_test),
                           h=365) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(temphumidity_forecast, series="ARIMA with Exogenous Variables",PI=FALSE) +
  ylab("Demand")

accuracy(temphumidity_forecast)

```

#### Discussion of ARIMA with 2 Exogenous Variables (Temp and Humidity)
This model did appear to fit the data well, with lower RMSE than previous model(s) (438.1585) and much lower MAPE as well (10.011). 

#### Full Forecast with both Exogenous Variables
```{r}
fullexog_arima <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
         xreg = cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
               temperature = ts_daily_temp, 
               humidity = ts_daily_humidity))
```

#### Forecasting Exogenous Variables for Full Data Forecast
In order to forecast using exogenous variables, you first have to forecast those exogenous variables. To start, rather than forecasting temperature and humidity, this forecast will just use the temperature values from the previous year during that time period. 
```{r}
#In order to forecast using exogenous variables you first have to forecast those exogenous variables. 

forecast_dailytemp <- daily_temp%>%
  filter(date > as.Date('2009-12-31') & date < ('2010-2-28'))
forecast_dailyhumid <- daily_humidity%>%
  filter(date > as.Date('2009-12-31') & date < ('2010-2-28'))


ts_temp_forfore <- msts(forecast_dailytemp$avg_t_ws1, 
                           seasonal.periods =c(7,365.25))

ts_humid_forfore <- msts(forecast_dailyhumid$rh_ws1, 
                           seasonal.periods =c(7,365.25))

```

```{r warning=FALSE, message=FALSE}
#Forecast Demand with Forecasted temperature and humidity
temphumidity_forecast_FULL <- forecast(tempandhumidity,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                temperature = ts_temp_forfore, 
                                humidity = ts_humid_forfore),
                           h=n_forecast) 

autoplot(temphumidity_forecast_FULL) + ylab("Demand")
accuracy(temphumidity_forecast_FULL)
```

### Export Model Results with Exogenous Variables
```{r}
temphumidexog_export <- data.frame(date=forecast_dates, load=temphumidity_forecast_FULL$mean)
write.csv(temphumidexog_export, "./Submissions/exog_humidtemp.csv", row.names = FALSE)
```

### Just Temperature
```{r}
temp_only<- auto.arima(ts_demand_train,
                  seasonal = FALSE, 
                  lambda = 0,
         xreg = as.matrix(fourier(ts_demand_train, 
                                          K=c(2,6)),
               temperature=ts_temp_train))
#summary(temp_only)

#Forecast 
temponly_forecast <- forecast(temp_only,
                           xreg=as.matrix(fourier(ts_demand_train, 
                                          K=c(2,6),h=365),
                                temperature=ts_temp_test),
                           h=365) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(temponly_forecast, series="ARIMA with Temperature",PI=FALSE) +
  ylab("Demand")

accuracy(temponly_forecast)
# An even higher RMSE than the temperature and humidity model. 
```

#### ARIMA with Temperature Exogenous Discussion
Although the plot appears to show that the lines match the observed data better, the MAPE and RMSE suggest that including temperature does not improve the model given that these values were higher than the RMSE/MAPE with both exogenous variables. Still, it is lower than other RMSE/MAPE values and so a full forecast can be performed. 

```{r warning=FALSE, message=FALSE}
fulltemp_arima <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
         xreg = cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
               temperature = ts_daily_temp))

#Forecast with 'forecasted' temperature, or temperature from previous year
temp_forecast_FULL <- forecast(fulltemp_arima,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                temperature = ts_temp_forfore),
                           h=n_forecast) 

autoplot(temp_forecast_FULL) + ylab("Demand")
accuracy(temp_forecast_FULL)
```

### Export Model Results with Exogenous Variables
```{r}
temp_export <- data.frame(date=forecast_dates, load=temp_forecast_FULL$mean)
write.csv(temp_export, "./Submissions/exog_temp.csv", row.names = FALSE)
```

### Just Humidity
```{r, warning = FALSE, message=FALSE}
humidity <- auto.arima(ts_demand_train,
                  seasonal = FALSE, 
                  lambda = 0,
         xreg = cbind(fourier(ts_demand_train, 
                                          K=c(2,6)),
               humidity=ts_humid_train))

#Forecast 
humidity_forecast <- forecast(humidity,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6)
                                          ,h=365 
                                         ),
                                humidity=ts_humid_test),
                           h=365) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(humidity_forecast, series="ARIMA with Humidity Exogenous Variable",PI=FALSE) +
  ylab("Demand")

accuracy(humidity_forecast)
```

#### ARIMA with Humidity Exogenous Discussion
The RMSE was 539.767 and the MAPE was 11.5054. 

```{r warning=FALSE, message=FALSE}
full_humidity_arima <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
         xreg = cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
               humidity = ts_daily_humidity))

#Forecast Demand with Forecasted humidity
humidity_forecast_FULL <- forecast(full_humidity_arima,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                humidity = ts_humid_forfore),
                           h=n_forecast) 

autoplot(humidity_forecast_FULL) + ylab("Demand")
accuracy(humidity_forecast_FULL)
```

### Export Model Results with Exogenous Variables
```{r}
humidity_export <- data.frame(date=forecast_dates, load=humidity_forecast_FULL$mean)
write.csv(humidity_export, "./Submissions/exog_humidity.csv", row.names = FALSE)
```


### Discussion of ARIMA with Exogenous Variables: 
Though none of the exogenous variable models scored highly based on the accuracy evaluations, the best model of the three included both exogenous variables. This will be selected to forecast using the full dataset. 




## ARIMA with Independent Variables
Just trying out random things here, but it looks bad, so not planning to upload it to Kaggle! We can remove before turning in the final RMD!
``` {r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
# Set up regressors: 
regressors <- data.frame(date = daily_demand$date, temp = daily_temp$avg_t_ws23, humidity = daily_humidity$rh_ws20)

regressors <- regressors %>%
  mutate(dow = wday(date),
         doy = yday(date)) %>%
  select(-c(date, dow, doy, humidity)) 
# , temp, humidity


X <- data.matrix(regressors)
X_train <- X[1:1826,]
X_test <- X[1827:2191,]

# First Model: 
ARIMA_reg_fit <- auto.arima(ts_demand_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=c(fourier(),X_train) # 2 weekly cycles, 6 annual cycles
                            # Luana - can combine the fourier coefficients and exogenous variables as a list here!

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_reg_forecast <- forecast(ARIMA_reg_fit,
                           xreg=X_test,
                           h=365
                           ) 

#Plot forecasting results
autoplot(ARIMA_reg_forecast) + ylab("Demand")

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(ARIMA_reg_forecast, series="ARIMA_Regressors",PI=FALSE) +
  ylab("Demand")

# Plot model and observed data during 2010
plot(ARIMA_reg_forecast$mean,
     xlab = "Date",
     ylab = "Demand",
     ylim = c(1000, 10000))
lines(ts_demand_test,
      type = "l",
      col = 4,
      lwd = 2)
legend(2010, 10000, legend=c("Forecasting Results", "Last Year of Data (2010)"),
       col=c("black",4), lwd=3, cex=0.8)
  # It's just terrible!

ARIMA_reg_scores <- accuracy(ARIMA_reg_forecast$mean, ts_demand_test)
checkresiduals(ARIMA_reg_forecast)

```


## TBATS Model 

```{r TBATS, echo=TRUE, message=FALSE, warning=FALSE}
# TBATS can take time to fit
TBATS_fit <- tbats(ts_demand_train)
TBATS_for <- forecast(TBATS_fit, h=365)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Active Power") 

#Plot model + observed data
autoplot(ts_demand_train) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Electricity Demand") 

# Plot model and observed data during 2010
plot(TBATS_for$mean,
     xlab = "Date",
     ylab = "Demand",
     ylim = c(1000, 10000))
lines(ts_demand_test,
      type = "l",
      col = 4,
      lwd = 2)
legend(2010, 10000, legend=c("Forecasting Results", "Last Year of Data (2010)"),
       col=c("black",4), lwd=3, cex=0.8)

TBATS_scores <- accuracy(TBATS_for$mean, ts_demand_test)
checkresiduals(TBATS_for)

```

### TBATS Interpretation
THE RMSE is 912.102 and the MAPE is 18.35. This model performed better than the ARIMA with fourier.

### Modeling Full Data - TBATS
After using the training dataset, re-running with the full dataset to forecast into 2011:
```{r}
# First Model: 
TBATS_fit_full <- tbats(ts_daily_demand)
TBATS_for_full <- forecast(TBATS_fit_full, h=n_forecast)

autoplot(TBATS_for_full) + ylab("Demand")

```

### Export Model Results - TBATs
```{r}
TBATS_export <- data.frame(date=forecast_dates, load=TBATS_for_full$mean)
write.csv(TBATS_export, "./Submissions/TBATS_v1.csv", row.names = FALSE)

```

## STL + ETS
```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(ts_demand_train,h=365)

#Plot foresting results
autoplot(ETS_fit) + ylab("Active Power")

#Plot model + observed data
autoplot(ts_demand_train) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Active Power")

ETS_scores <- accuracy(ETS_fit$mean, ts_demand_test)
checkresiduals(ETS_fit)
```

### STL + ETS Interpretation 
The RMSE is 1231.962 and the MAPE is 33.30224, both higher than the original ARIMA_Fourier.  

### Full Data Modeling - STL+ETS
```{r}
#Fit and forecast STL + ETS model to data
ETS_fit_full <-  stlf(ts_daily_demand,h=n_forecast)

#Plot foresting results
autoplot(ETS_fit) + ylab("Active Power")

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(ETS_fit_full, series="STL + ETS",PI=FALSE) +
  ylab("Active Power")

```

### Export Model Results - ETS
```{r}
ETS_STL_export <- data.frame(date=forecast_dates, load=ETS_fit_full$mean)
write.csv(ETS_STL_export, file="./Submissions/ETS_STL_v1.csv", row.names=FALSE)

```

## Neural Network Model 1 
```{r}
# Fit Model on Training Data
NN_fit <- nnetar(ts_demand_train,
                 p=1,
                 P=0,
                 xreg=fourier(ts_demand_train, K=c(3.5,2)))

# Forecast on Testing Data
NN_for <- forecast(NN_fit, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(3.5,2),h=365))

#Plot foresting results
autoplot(NN_for) +
  ylab("Electricity Demand") 

#Plot model + observed data
autoplot(ts_demand_train) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Plot model and observed data during 2010
plot(NN_for$mean,
     xlab = "Date",
     ylab = "Demand",
     ylim = c(1000, 10000))
lines(ts_demand_test,
      type = "l",
      col = 4,
      lwd = 2)
legend(2010, 10000, legend=c("Forecasting Results", "Last Year of Data (2010)"),
       col=c("black",4), lwd=3, cex=0.8)

accuracy(NN_for$mean, ts_demand_test)
NN_scores <- accuracy(NN_for$mean, ts_demand_test)
checkresiduals(NN_for)
```

### NN Interpretation
THE RMSE is 1000.898 and the MAPE is 16.67504. The MAPE appears to be the lowest among models, but RMSE is in the middle of the range.

### Neural Network Output Submission 1.5
I called this 1.5 because our initial neural network output v1 was accidentally the fitted data of an STL-ETS model. So I just tweaked it to make sure it was exporting the right model! The original code is at the bottom where I am saving code I don't want to delete but just want to streamline the RMD. 
```{r}
#Fit and forecast full neural network
NN_fit_full <-  nnetar(ts_daily_demand,
                 p=1,
                 P=0,
                 xreg=fourier(ts_daily_demand, K=c(3.5,2)))

NN_full_for <- forecast(NN_fit_full, 
                   h=365,
                   xreg=fourier(ts_daily_demand, 
                                          K=c(3.5,2),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for)

```

### Export Model Results - NN
```{r}
NN_export15 <- data.frame(date=forecast_dates, load=NN_full_for$mean)
write.csv(NN_export15, file="./Submissions/NN_v1p5.csv", row.names=FALSE)

```


## Neural Network Model 2
```{r}
# Fit Model on Training Data
NN_fit2 <- nnetar(ts_demand_train,
                 p=2,
                 P=0,
                 xreg=fourier(ts_demand_train, K=c(2,6)))

# Forecast on Testing Data
NN_for2 <- forecast(NN_fit2, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=365))

#Plot foresting results
autoplot(NN_for2) +
  ylab("Electricity Demand") 

#Plot model + observed data
autoplot(ts_demand_train) +
  autolayer(NN_for2, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Plot model and observed data during 2010
plot(NN_for2$mean,
     xlab = "Date",
     ylab = "Demand",
     ylim = c(1000, 10000))
lines(ts_demand_test,
      type = "l",
      col = 4,
      lwd = 2)
legend(2010, 10000, legend=c("Forecasting Results", "Last Year of Data (2010)"),
       col=c("black",4), lwd=3, cex=0.8)

accuracy(NN_for2$mean, ts_demand_test)
NN2_scores <- accuracy(NN_for2$mean, ts_demand_test)
checkresiduals(NN_for2)
```

### NN2 Interpretation
THE RMSE is 1092.151 and the MAPE is 17.423. The MAPE is higher than the other neural network model, but still relatively low. 

### Full Data Modeling - Neural Network 2
```{r}
#Fit and forecast full neural network
NN2_fit_full <-  nnetar(ts_daily_demand,
                 p=2,
                 P=0,
                 xreg=fourier(ts_daily_demand, K=c(2,6)))

NN_full_for2 <- forecast(NN2_fit_full, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for2, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for2)

```

### Export Model Results - NN2
```{r}
NN2_export <- data.frame(date=forecast_dates, load=NN_full_for2$mean)
write.csv(NN2_export, file="./Submissions/NN_v2.csv", row.names=FALSE)

```

## Compare All Models 
```{r}
scores <- as.data.frame(rbind(ARIMA_four_scores, ARIMA_reg_scores, TBATS_scores, ETS_scores, NN_scores))
row.names(scores) <- c("ARIMA_for", "ARIMA_reg","TBATS", "ETS", "NN")

```

### ARIMA with exogenous variables 
``` {r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
# First Model: 
ARIMA_Fourier_fit2 <- auto.arima(ts_demand_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_demand_train,
                                          K=c(2,2)) # 2 weekly cycles, 2 annual cycles
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Fourier_forecast <- forecast(ARIMA_Fourier_fit2,
                           xreg=fourier(ts_demand_train,
                                        K=c(2,2),
                                        h=365),
                           h=365
                           ) 

#Plot foresting results
autoplot(ARIMA_Fourier_forecast) + ylab("Demand")

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(ARIMA_Fourier_forecast, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Demand")

# Plot model and observed data during 2010
plot(ARIMA_Fourier_forecast$mean,
     xlab = "Date",
     ylab = "Demand",
     ylim = c(1000, 10000))
lines(ts_demand_test,
      type = "l",
      col = 4,
      lwd = 2)
legend(2010, 10000, legend=c("Forecasting Results", "Last Year of Data (2010)"),
       col=c("black",4), lwd=3, cex=0.8)

accuracy(ARIMA_Fourier_fit)
```



## Saving Code
## I started forecasting temperature then stopped: 
```{r}

# Decompose 
ts_daily_temp %>% mstl() %>%
  autoplot()

temp_arima <- auto.arima(ts_daily_temp, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_daily_temp, 
                                          K=c(2,2)) # 2 weekly cycles, 6 annual cycles
                             )


temp_forecast <- forecast(temp_arima,
                           xreg=fourier(ts_daily_temp,
                                        K=c(2,2),
                                        h=n_forecast),
                           h=n_forecast
                           ) 

autoplot(temp_forecast) + ylab("Demand")
```


Pasted Code from Original Neural Network export: 

## Full Data Modeling - Neural Network
```{r}
#Fit and forecast STL + ETS model to data
NN_fit_full <-  stlf(ts_daily_demand,h=n_forecast)
#This STLF code runs the STL+ETS model, so I don't know if we've yet exported the full model results from the first Neural Network! I just did so, below -- I could be wrong, so I didn't delete this! 

#Plot foresting results
autoplot(NN_fit_full) + ylab("Active Power")

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_fit_full, series="STL + ETS",PI=FALSE) +
  ylab("Active Power")

```

## Export Model Results - NN --> 
```{r}
NN_export <- data.frame(date=forecast_dates, load=NN_fit_full$mean)
write.csv(NN_export, file="./Submissions/NN_v1.csv", row.names=FALSE)

```






I saw this pasted from the in-class exercise -- I didn't want to delete it all together so I just put it down here to streamline


```{r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
#Fit arima model with fourier terms as exogenous regressors
# seasonal = FALSE is the same as P=D=Q=0
# play with K by changing it to K=c(2,2), K=c(2,4), K=c(2,6), etc. The higher teh K the longer it will take to converge, because R will try more models.


ARIMA_Fourier_fit <- auto.arima(ts_demand_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_demand_train, 
                                          K=c(2,2)) # 2 weekly cycles, 2 annual cycles
                             )

ARIMA_Four_fit <- auto.arima(ts_act_power_daily_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_act_power_daily_train, 
                                          K=c(2,12))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(ts_act_power_daily_train,
                                        K=c(2,12),
                                        h=365),
                           h=365
                           ) 

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Active Power")

#Plot model + observed data
autoplot(ts_act_power_daily) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Active Power")

```

