---
title: "Wrangling Competition Data"
author: "Katherine Burley & Maggie O'Shea"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Load Libraries
```{r message = FALSE, warning=FALSE}
rm(list=ls())
library(tidyverse)
library(readxl)
library(writexl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(kableExtra)
```

## Import the Dataset
```{r warning=FALSE}
hourly_demand <- read_excel("./Data/load.xlsx")
hourly_temp<- read_excel("./Data/temperature.xlsx")
hourly_humidity<- read_excel("./Data/relative_humidity.xlsx")
submission_template <- read_excel("./Data/submission_template.xlsx")
```

## Wrangle/Process the Dataset
```{r warning=FALSE}
# Demand
daily_demand <- hourly_demand %>%
  pivot_longer(cols=h1:h24, names_to = "hour", values_to = "demand") %>%
  filter(!is.na(demand)) %>% # only a handful of missing hourly obs. Let's try removing them
  group_by(date, meter_id) %>%
  summarise(daily_average = mean(demand))

# Temperature
daily_temp <- hourly_temp%>%
  group_by(date)%>%
  summarize(avg_t_ws1 = mean(t_ws1),
            avg_t_ws2 = mean(t_ws2),
            avg_t_ws3 = mean(t_ws3),
            avg_t_ws4 = mean(t_ws4),
            avg_t_ws5 = mean(t_ws5),
            avg_t_ws6 = mean(t_ws6),
            avg_t_ws7 = mean(t_ws7),
            avg_t_ws8 = mean(t_ws8),
            avg_t_ws9 = mean(t_ws9),
            avg_t_ws10 = mean(t_ws10),
            avg_t_ws11 = mean(t_ws11),
            avg_t_ws12 = mean(t_ws12),
            avg_t_ws13 = mean(t_ws13),
            avg_t_ws13 = mean(t_ws13),
            avg_t_ws14 = mean(t_ws14),
            avg_t_ws15 = mean(t_ws15),
            avg_t_ws16 = mean(t_ws16),
            avg_t_ws17 = mean(t_ws17),
            avg_t_ws18 = mean(t_ws18),
            avg_t_ws19 = mean(t_ws19),
            avg_t_ws20 = mean(t_ws20),
            avg_t_ws21 = mean(t_ws21),
            avg_t_ws22 = mean(t_ws22),
            avg_t_ws23 = mean(t_ws23),
            avg_t_ws24 = mean(t_ws24),
            avg_t_ws25 = mean(t_ws25),
            avg_t_ws26 = mean(t_ws26),
            avg_t_ws27 = mean(t_ws27),
            avg_t_ws28 = mean(t_ws28))

# Humidity
daily_humidity <- hourly_humidity%>%
  pivot_longer(cols = -c(hr, date), names_to = "rh", values_to = "humidity")%>%
  group_by(date, rh)%>%
  summarize(avg_humidity = mean(humidity))%>%
  pivot_wider(values_from = "avg_humidity", names_from = "rh")

# Extract forecast dates from the submission template
forecast_dates <- submission_template$date
rm(cor_mat, daily_all, daily_forcor, daily_forcor_mat, submission_template)
  
```

## Create a Time Series Object
```{r warning=FALSE}
# Create ts object
ts_daily_demand <- msts(daily_demand$daily_average, 
                           seasonal.periods =c(7, 365.25),
                           start=c(2005,01,01))

ts_daily_humidity <- msts(daily_humidity$rh_ws20, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))
ts_daily_temp <- msts(daily_temp$avg_t_ws23, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))
  # Using temperature and humidity vars with highest correlation to demand - see next chunk!

# Subset training data
ts_demand_train <- subset(ts_daily_demand,
                          end = length(ts_daily_demand)-365)
ts_humid_train <- subset(ts_daily_humidity,
                          end = length(ts_daily_demand)-365)
ts_temp_train <- subset(ts_daily_temp,
                          end = length(ts_daily_demand)-365)

# Subset testing data
ts_demand_test <- subset(ts_daily_demand,
                         start = length(ts_daily_demand)-365)
ts_demand_test <- subset(ts_demand_test,
                         end = length(ts_demand_test)-1) 

ts_temp_test <- subset(ts_daily_temp,
                          start = length(ts_daily_demand)-365)
ts_temp_test <- subset(ts_temp_test,
                         end = length(ts_temp_test)-1) 

ts_humid_test <- subset(ts_daily_humidity,
                          start = length(ts_daily_demand)-365)
ts_humid_test <- subset(ts_humid_test,
                         end = length(ts_humid_test)-1) 
```

## Decompose and Check the ACF + PACF

```{r warning=FALSE}
# Decompose 
ts_daily_demand %>% mstl() %>%
  autoplot()

decompose <- data.frame(mstl(ts_daily_demand))
ts_remainder <- ts(decompose$Remainder)

## Look for correlations between humidity, temp, and demand/de-seasoned demand
deseasoned <- data.frame(date = daily_demand$date, deseas_demand = ts_remainder)
daily_all <- daily_demand %>%
  left_join(deseasoned, by="date") %>%
  left_join(daily_humidity, by="date") %>%
  left_join(daily_temp, by="date") %>%
  ungroup()

daily_forcor <- daily_all %>%
  select(-c(date, meter_id))

daily_forcor_mat <- data.matrix(daily_forcor)

cor_mat <- cor(daily_forcor, use="complete")
cor_results <- data.frame(corr = cor_mat[,1:2])
cor_results <- cor_results %>%
  mutate(corr.daily_average = abs(corr.daily_average),
         corr.deseas_demand = abs(corr.deseas_demand))
  # none of the humidity or temp vars have a very strong correlation with demand or deseasoned demand...

# Plot the time series over time, ACF and PACF
par(mfrow=c(1,3), mar=c(6, 4, 4, 2))
plot(ts_remainder, main="Remainder Time Series", ylab = "Demand")
Acf(ts_remainder,lag.max=40, main="ACF") 
Pacf(ts_remainder,lag.max=40, main="PACF")
  # Looks like an AR process.
  # There is also still some seasonality visible in the PACF.

```

## Temperature and Humidity Predictions for Models with Exogenous Variables

### Use last year's values
In order to forecast using exogenous variables, you first have to forecast those exogenous variables. To start, rather than forecasting temperature and humidity, this forecast will just use the temperature values from the previous year during that time period.
```{r warning=FALSE}
#In order to forecast using exogenous variables you first have to forecast those exogenous variables. 
forecast_dailytemp <- daily_temp%>%
  filter(date > as.Date('2009-12-31') & date < ('2010-2-28'))
forecast_dailyhumid <- daily_humidity%>%
  filter(date > as.Date('2009-12-31') & date < ('2010-2-28'))


ts_temp_forfore <- msts(forecast_dailytemp$avg_t_ws1,
                        seasonal.periods =c(7,365.25),
                        start=c(2011,01,01))

ts_humid_forfore <- msts(forecast_dailyhumid$rh_ws1,
                         seasonal.periods =c(7,365.25),
                         start=c(2011,01,01))

```

### Use average daily values
As an alternative option, we will use average daily values of temperature and humidity from the observable data as the forecast for 2011.
```{r warning=FALSE}
# Use average daily values as forecast for temperature and humidity
temp_for <- daily_temp %>%
  select(date, avg_t_ws23) %>%
  mutate(doy = yday(date)) %>%
  group_by(doy) %>%
  summarise(avg_temp = mean(avg_t_ws23)) %>%
  ungroup() %>%
  filter(doy<=59)

ts_temp_for <- msts(temp_for$avg_temp,
                    seasonal.periods =c(7,365.25),
                    start=c(2011,01,01))

humid_for <- daily_humidity %>%
  select(date, rh_ws20) %>%
  mutate(doy = yday(date)) %>%
  group_by(doy) %>%
  summarise(avg_humid = mean(rh_ws20)) %>%
  ungroup() %>%
  filter(doy<=59)

ts_humid_for <- msts(temp_for$avg_temp,
                    seasonal.periods =c(7,365.25),
                    start=c(2011,01,01))

# Temperature
autoplot(ts_temp_train, series="Train") +
  autolayer(ts_temp_test, series="Test") +
  autolayer(ts_temp_forfore, series="Forecast1") +
  autolayer(ts_temp_for, series="Forecast 2") +
  ylab("Temperature")

# Humidity
autoplot(ts_humid_train, series="Train") +
  autolayer(ts_humid_test, series="Test") +
  autolayer(ts_humid_forfore, series="Forecast1") +
  autolayer(ts_humid_for, series="Forecast 2") +
  ylab("Humidity")

# Will use Forecast 2 in the models.

```

# Fit Models to your Data

Fit models to your dataset considering the period Jan 1st 2005 to Dec 31st 2009. First, we will consider an ARIMA with Fourier terms to account for multiple seasonalities in our data. For this specification, we assume that there are two sine and cosine terms during the weekly season and six sine and cosine terms during the annual season, after testing different combinations of K.  

## ARIMA with Fourier Terms Model 
``` {r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
# First Model: 
ARIMA_Fourier_fit <- auto.arima(ts_demand_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_demand_train, 
                                          K=c(2,6)) # 2 weekly cycles, 6 annual cycles
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Fourier_forecast <- forecast(ARIMA_Fourier_fit,
                           xreg=fourier(ts_demand_train,
                                        K=c(2,6),
                                        h=365),
                           h=365
                           ) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(ARIMA_Fourier_forecast, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Demand")

# Plot residuals
checkresiduals(ARIMA_Fourier_forecast)

# Calculate Scores
ARIMA_four_scores <- accuracy(ARIMA_Fourier_forecast$mean, ts_demand_test)

```

### ARIMA with Fourier - Training Data Interpretation
THE RMSE is 951.8972 and the MAPE is 23.426. 

### Modeling Full Data - ARIMA with Fourier
After using the training dataset, re-running with the full dataset to forecast into 2011:
```{r}
# First Model: 
ARIMA_Fourier_full <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_daily_demand, 
                                          K=c(2,2)) # 2 weekly cycles, 2 annual cycles
                             )
n_forecast = 31+28 # January and February - 59 days

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Fourier_FULLforecast <- forecast(ARIMA_Fourier_full,
                           xreg=fourier(ts_daily_demand,
                                        K=c(2,2),
                                        h=n_forecast),
                           h=n_forecast
                           ) 

# Plot Forecast
autoplot(ARIMA_Fourier_FULLforecast) + ylab("Demand")

```

### Export Model Results
```{r}
ARIMA_fourier_export <- data.frame(date=forecast_dates, load=ARIMA_Fourier_FULLforecast$mean)
# write.csv(ARIMA_fourier_export, "./Submissions/ARIMA_fourier_v3.csv", row.names = FALSE)
```

## ARIMA with Exogenous Variables

### Both Temperature and Humidity 
```{r warnings = FALSE}
tempandhumidity <- auto.arima(ts_demand_train,
                  seasonal = FALSE, 
                  lambda = 0,
         xreg = cbind(fourier(ts_demand_train, 
                                          K=c(2,6)),
               humidity=ts_humid_train,
               temperature=ts_temp_train))



#Forecast 
temphumidity_forecast <- forecast(tempandhumidity,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),h=365),
                                humidity=ts_humid_test,
                                temperature=ts_temp_test),
                           h=365) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(temphumidity_forecast, series="ARIMA with Exog Vars",PI=FALSE) +
  ylab("Demand")

# Plot residuals
checkresiduals(temphumidity_forecast)

# Calculate Scores
Exog1_scores <- accuracy(temphumidity_forecast$mean, ts_demand_test)

```

#### Discussion of ARIMA with 2 Exogenous Variables (Temp and Humidity)
This model did appear to fit the data well, with lower RMSE than previous model(s) (404.31) and much lower MAPE as well (9.35). 

#### Full Forecast with both Exogenous Variables
```{r warnings = FALSE}
fullexog_arima <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
         xreg = cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
                      temperature = ts_daily_temp,
                      humidity = ts_daily_humidity))

#Forecast Demand with Forecasted temperature and humidity
temphumidity_forecast_FULL <- forecast(tempandhumidity,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                temperature = ts_temp_for, 
                                humidity = ts_humid_for),
                           h=n_forecast) 

autoplot(temphumidity_forecast_FULL) + ylab("Demand")
```

#### Export Model Results with Exogenous Variables
```{r warnings = FALSE}
temphumidexog_export <- data.frame(date=forecast_dates, load=temphumidity_forecast_FULL$mean)
write.csv(temphumidexog_export, "./Submissions/exog_humidtemp_v3.csv", row.names = FALSE)
```

### Just Temperature Exogenous Variable
```{r warnings = FALSE}
temp_only<- auto.arima(ts_demand_train,
                  seasonal = FALSE, 
                  lambda = 0,
         xreg = as.matrix(fourier(ts_demand_train, 
                                          K=c(2,6)),
               temperature=ts_temp_train))
#summary(temp_only)

#Forecast 
temponly_forecast <- forecast(temp_only,
                           xreg=as.matrix(fourier(ts_demand_train, 
                                          K=c(2,6),h=365),
                                temperature=ts_temp_test),
                           h=365) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(temponly_forecast, series="ARIMA with Temperature",PI=FALSE) +
  ylab("Demand")

# Plot residuals
checkresiduals(temponly_forecast)

# Calculate scores
accuracy(temponly_forecast)
Exog2_scores <- accuracy(temponly_forecast$mean, ts_demand_test)

# A higher RMSE than the temperature and humidity model. 
```

#### ARIMA with Temperature Exogenous Discussion
Although the plot appears to show that the lines match the observed data better, the MAPE and RMSE suggest that including temperature does not improve the model given that these values were higher than the RMSE/MAPE with both exogenous variables. Still, it is lower than other RMSE/MAPE values and so a full forecast can be performed. 

#### Full Dataset Humidity Exogenous Forecast

```{r warnings = FALSE}
fulltemp_arima <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
         xreg = cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
               temperature = ts_daily_temp))

#Forecast with 'forecasted' temperature
temp_forecast_FULL <- forecast(fulltemp_arima,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                temperature = ts_temp_for),
                           h=n_forecast) 

autoplot(temp_forecast_FULL) + ylab("Demand")
```

#### Export Model Results with Temperature Variables
```{r warnings = FALSE}
temp_export <- data.frame(date=forecast_dates, load=temp_forecast_FULL$mean)
write.csv(temp_export, "./Submissions/exog_temp_v3.csv", row.names = FALSE)
```

### Just Humidity Exogenous Variable
```{r warnings = FALSE}
humidity <- auto.arima(ts_demand_train,
                  seasonal = FALSE, 
                  lambda = 0,
         xreg = cbind(fourier(ts_demand_train, 
                                          K=c(2,6)),
               humidity=ts_humid_train))

#Forecast 
humidity_forecast <- forecast(humidity,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6)
                                          ,h=365 
                                         ),
                                humidity=ts_humid_test),
                           h=365) 

#Plot model + all observed data
autoplot(ts_daily_demand) +
  autolayer(humidity_forecast, series="ARIMA with Humidity Exogenous Variable",PI=FALSE) +
  ylab("Demand")

# Calculate scores
accuracy(humidity_forecast)
Exog3_scores <- accuracy(humidity_forecast$mean, ts_demand_test)
```

#### ARIMA with Humidity Exogenous Discussion
The RMSE was 499.3725 and the MAPE was 11.26524. 

#### Full Dataset Humidity Exogenous Forecast
```{r warning=FALSE, message=FALSE}
full_humidity_arima <- auto.arima(ts_daily_demand, 
                             seasonal=FALSE, 
                             lambda=0,
         xreg = cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
               humidity = ts_daily_humidity))

#Forecast Demand with Forecasted humidity
humidity_forecast_FULL <- forecast(full_humidity_arima,
                           xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                humidity = ts_humid_for),
                           h=n_forecast) 

autoplot(humidity_forecast_FULL) + ylab("Demand")
```

#### Export Model Results with Exogenous Variables
```{r}
humidity_export <- data.frame(date=forecast_dates, load=humidity_forecast_FULL$mean)
# write.csv(humidity_export, "./Submissions/exog_humidity_v3.csv", row.names = FALSE)
```

## TBATS
```{r TBATS, echo=TRUE, message=FALSE, warning=FALSE}
# TBATS can take time to fit
TBATS_fit <- tbats(ts_demand_train)
TBATS_for <- forecast(TBATS_fit, h=365)

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Electricity Demand") 

# Plot residuals
checkresiduals(TBATS_for)

# Calculate Scores
TBATS_scores <- accuracy(TBATS_for$mean, ts_demand_test)

```

### TBATS Interpretation
THE RMSE is 912.102 and the MAPE is 18.35. This model performed better than the ARIMA with fourier.

### Modeling Full Data - TBATS
After using the training dataset, re-running with the full dataset to forecast into 2011:
```{r}
# First Model: 
TBATS_fit_full <- tbats(ts_daily_demand)
TBATS_for_full <- forecast(TBATS_fit_full, h=n_forecast)

autoplot(TBATS_for_full) + ylab("Demand")

```

### Export Model Results - TBATs
```{r}
TBATS_export <- data.frame(date=forecast_dates, load=TBATS_for_full$mean)
# write.csv(TBATS_export, "./Submissions/TBATS_v1.csv", row.names = FALSE)

```

## STL + ETS
```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(ts_demand_train,h=365)

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Active Power")

# Plot Residuals
checkresiduals(ETS_fit)

# Calculate Scores
ETS_scores <- accuracy(ETS_fit$mean, ts_demand_test)
```

### STL + ETS Interpretation 
The RMSE is 1231.962 and the MAPE is 33.30224, both higher than the original ARIMA_Fourier.  

### Full Data Modeling - STL+ETS
```{r}
#Fit and forecast STL + ETS model to data
ETS_fit_full <-  stlf(ts_daily_demand,h=n_forecast)

#Plot foresting results
autoplot(ETS_fit) + ylab("Active Power")

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(ETS_fit_full, series="STL + ETS",PI=FALSE) +
  ylab("Active Power")

```

### Export Model Results - ETS
```{r}
ETS_STL_export <- data.frame(date=forecast_dates, load=ETS_fit_full$mean)
# write.csv(ETS_STL_export, file="./Submissions/ETS_STL_v1.csv", row.names=FALSE)

```

## Neural Network Model 1 
```{r}
# Fit Model on Training Data
NN_fit <- nnetar(ts_demand_train,
                 p=1,
                 P=0,
                 xreg=fourier(ts_demand_train, K=c(3.5,2)))

# Forecast on Testing Data
NN_for <- forecast(NN_fit, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(3.5,2),h=365))

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Plot residuals
checkresiduals(NN_for)

# Calculate scores
accuracy(NN_for$mean, ts_demand_test)
NN_scores <- accuracy(NN_for$mean, ts_demand_test)

```

### NN Interpretation
THE RMSE is 997.6037 and the MAPE is 16.52452. The MAPE appears to be the lowest among models, but RMSE is in the middle of the range.

### Neural Network Output Submission 1.5
I called this 1.5 because our initial neural network output v1 was accidentally the fitted data of an STL-ETS model. So I just tweaked it to make sure it was exporting the right model! The original code is at the bottom where I am saving code I don't want to delete but just want to streamline the RMD. 
```{r}
#Fit and forecast full neural network
NN_fit_full <-  nnetar(ts_daily_demand,
                 p=1,
                 P=0,
                 xreg=fourier(ts_daily_demand, K=c(3.5,2)))

NN_full_for <- forecast(NN_fit_full, 
                   h=365,
                   xreg=fourier(ts_daily_demand, 
                                          K=c(3.5,2),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for)

```

### Export Model Results - NN
```{r}
NN_export15 <- data.frame(date=forecast_dates, load=NN_full_for$mean)
# write.csv(NN_export15, file="./Submissions/NN_v1p5.csv", row.names=FALSE)

```

## Neural Network Model 2
```{r}
# Fit Model on Training Data
NN_fit2 <- nnetar(ts_demand_train,
                 p=2,
                 P=0,
                 xreg=fourier(ts_demand_train, K=c(2,6)))

# Forecast on Testing Data
NN_for2 <- forecast(NN_fit2, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=365))


#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for2, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Plot residuals
checkresiduals(NN_for2)

# Calculate scores
accuracy(NN_for2$mean, ts_demand_test)
NN2_scores <- accuracy(NN_for2$mean, ts_demand_test)

```

### NN2 Interpretation
THE RMSE is 1088.129 and the MAPE is 17.39624 The MAPE is higher than the other neural network model, but still relatively low. 

### Full Data Modeling - Neural Network 2
```{r}
#Fit and forecast full neural network
NN2_fit_full <-  nnetar(ts_daily_demand,
                 p=2,
                 P=0,
                 xreg=fourier(ts_daily_demand, K=c(2,6)))

NN_full_for2 <- forecast(NN2_fit_full, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for2, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for2)

```

### Export Model Results - NN2
```{r}
NN2_export <- data.frame(date=forecast_dates, load=NN_full_for2$mean)
# write.csv(NN2_export, file="./Submissions/NN_v2.csv", row.names=FALSE)

```

## Neural Network Model 3
```{r}
# Fit Model on Training Data
NN_fit3 <- nnetar(ts_demand_train,
                 p=1,
                 P=1,
                 xreg=fourier(ts_demand_train, K=c(2,6)))

# Forecast on Testing Data
NN_for3 <- forecast(NN_fit3, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=365))


#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for3, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Calculate scores
accuracy(NN_for3$mean, ts_demand_test)
NN3_scores <- accuracy(NN_for3$mean, ts_demand_test)
```

### NN3 Interpretation
THE RMSE is 1158.751 and the MAPE is 18.84508. 

### Full Data Modeling - Neural Network 3
```{r}
#Fit and forecast full neural network
NN3_fit_full <-  nnetar(ts_daily_demand,
                 p=1,
                 P=1,
                 xreg=fourier(ts_daily_demand, K=c(2,6)))

NN_full_for3 <- forecast(NN3_fit_full, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for3, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for3)

```

### Export Model Results - NN3
```{r}
NN3_export <- data.frame(date=forecast_dates, load=NN_full_for3$mean)
# write.csv(NN3_export, file="./Submissions/NN_v3.csv", row.names=FALSE)

```

## Neural Network Model 4
```{r}
# Fit Model on Training Data
NN_fit4 <- nnetar(ts_demand_train,
                 p=1,
                 P=2,
                 xreg=fourier(ts_demand_train, K=c(2,6)))

# Forecast on Testing Data
NN_for4 <- forecast(NN_fit4, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=365))

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for4, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_for4$mean, ts_demand_test)
NN4_scores <- accuracy(NN_for4$mean, ts_demand_test)
```

### NN5 Interpretation
THE RMSE is 1119.653 and the MAPE is 19.45936. 

### Full Data Modeling - Neural Network 4
```{r}
#Fit and forecast full neural network
NN4_fit_full <-  nnetar(ts_daily_demand,
                 p=1,
                 P=2,
                 xreg=fourier(ts_daily_demand, K=c(2,6)))

NN_full_for4 <- forecast(NN4_fit_full, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for4, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for4)

```

### Export Model Results - NN4
```{r}
NN4_export <- data.frame(date=forecast_dates, load=NN_full_for4$mean)
# write.csv(NN4_export, file="./Submissions/NN_v4.csv", row.names=FALSE)

```

## Neural Network Model 5
```{r}
# Fit Model on Training Data
NN_fit5 <- nnetar(ts_demand_train,
                 p=2,
                 P=1,
                 xreg=fourier(ts_demand_train, K=c(2,6)))

# Forecast on Testing Data
NN_for5 <- forecast(NN_fit5, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=365))

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for5, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Calculate scores
accuracy(NN_for5$mean, ts_demand_test)
NN5_scores <- accuracy(NN_for5$mean, ts_demand_test)
```

### NN5 Interpretation
THE RMSE is 1095.736  and the MAPE is 18.15347.  

### Full Data Modeling - Neural Network 5
```{r}
#Fit and forecast full neural network
NN5_fit_full <-  nnetar(ts_daily_demand,
                 p=2,
                 P=1,
                 xreg=fourier(ts_daily_demand, K=c(2,6)))

NN_full_for5 <- forecast(NN5_fit_full, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for5, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for5)

```

### Export Model Results - NN5
```{r}
NN5_export <- data.frame(date=forecast_dates, load=NN_full_for5$mean)
# write.csv(NN5_export, file="./Submissions/NN_v5.csv", row.names=FALSE)

```

## Neural Network Model 6
```{r}
# Fit Model on Training Data
NN_fit6 <- nnetar(ts_demand_train,
                 p=2,
                 P=2,
                 xreg=fourier(ts_demand_train, K=c(2,6)))

# Forecast on Testing Data
NN_for6 <- forecast(NN_fit6, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=365))


#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for6, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

# Calculate Scores
accuracy(NN_for6$mean, ts_demand_test)
NN6_scores <- accuracy(NN_for6$mean, ts_demand_test)
```

### NN6 Interpretation
THE RMSE is 1055.742 and the MAPE is 17.88271. 
Across all NN models, V6 (p=2, P=2) scored the best on Kaggle, but V1 (p=2,P=0) scores the best in terms of RMSE and MAPE.

### Full Data Modeling - Neural Network 6
```{r}
#Fit and forecast full neural network
NN6_fit_full <-  nnetar(ts_daily_demand,
                 p=2,
                 P=2,
                 xreg=fourier(ts_daily_demand, K=c(2,6)))

NN_full_for6 <- forecast(NN6_fit_full, 
                   h=365,
                   xreg=fourier(ts_demand_train, 
                                          K=c(2,6),h=59))
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for6, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for6)

```

### Export Model Results - NN6
```{r}
NN6_export <- data.frame(date=forecast_dates, load=NN_full_for6$mean)
# write.csv(NN6_export, file="./Submissions/NN_v6.csv", row.names=FALSE)

```

## Neural Network Model 7 - Include Exogenous Variables
```{r}
# Fit Model on Training Data
NN_fit7 <- nnetar(ts_demand_train,
                 p=2,
                 P=2,
                 xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6)),
                            humidity=ts_humid_train,
                            temperature=ts_temp_train))

# Forecast on Testing Data
NN_for7 <- forecast(NN_fit7, 
                   h=365,
                   xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),h=365),
                                humidity=ts_humid_test,
                                temperature=ts_temp_test))

#Plot model + observed data
autoplot(ts_daily_demand) +
  autolayer(NN_for7, series="Neural Network + Exog",PI=FALSE)+
  ylab("Electricity Demand") 

# Calculate scores
accuracy(NN_for7$mean, ts_demand_test)
NNexog_scores <- accuracy(NN_for7$mean, ts_demand_test)
```

### NN7 Interpretation
THE RMSE is 599.2527 and the MAPE is 12.05602.  

### Full Data Modeling - Neural Network 7
```{r}
#Fit and forecast full neural network
NN7_fit_full <-  nnetar(ts_daily_demand,
                 p=2,
                 P=2,
                 xreg=cbind(fourier(ts_daily_demand, 
                                          K=c(2,6)),
                            temperature = ts_daily_temp,
                            humidity = ts_daily_humidity))

NN_full_for7 <- forecast(NN7_fit_full, 
                   xreg=cbind(fourier(ts_demand_train, 
                                          K=c(2,6),
                                          h=n_forecast),
                                temperature = ts_temp_for, 
                                humidity = ts_humid_for),
                   h=n_forecast)
#Plot foresting 
autoplot(ts_daily_demand) +
  autolayer(NN_full_for7, series="Neural Network",PI=FALSE)+
  ylab("Electricity Demand") 

accuracy(NN_full_for7)

```

### Export Model Results - NN7
```{r}
NN7_export <- data.frame(date=forecast_dates, load=NN_full_for7$mean)
# write.csv(NN7_export, file="./Submissions/NN_v7.csv", row.names=FALSE)

```


## Compare All Models 
```{r Best Model}
scores <- as.data.frame(rbind(ARIMA_four_scores, Exog1_scores, Exog2_scores, Exog3_scores, 
                              TBATS_scores, ETS_scores, NN_scores, NN6_scores, NNexog_scores))

row.names(scores) <- c("ARIMA+Fourier", "ARIMA+Fourier+Temp+Humid", "ARIMA+Fourier+Temp", 
                       "ARIMA+Fourier+Humid", "TBATS", "ETS", "NN(p=2,P=2)", "NN(p=2,P=0)", "NN+Temp+Humid")

# Which model has the best RMSE?
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))  
```

```{r Table}
# Table of Scores
kbl(scores,
    caption = "Forecast Accuracy for Daily Demand",
    digits = array(9,ncol(scores)),
    table.attr = "style = \"color: black;\"") %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") %>%
  kable_styling() %>%
  row_spec(row=which.min(scores[,"RMSE"]), bold = T, color = "white", background = "cornflowerblue")

```

```{r Visuals}
# Plot Forecasting Results - All Data
autoplot(ts_daily_demand) +
  autolayer(ARIMA_Fourier_forecast, series="ARIMA + FOURIER",PI=FALSE) +
  autolayer(temphumidity_forecast, series="ARIMA + Fourier + Temp + Humidity",PI=FALSE) +
  autolayer(temponly_forecast, series="ARIMA + Fourier + Temp",PI=FALSE) +
  autolayer(humidity_forecast, series="ARIMA + Fourier + Humidity",PI=FALSE) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE)+
  autolayer(NN_for, series="NN(p=2,P=2)",PI=FALSE)+
  autolayer(NN_for6, series="NN(p=2,P=0)",PI=FALSE)+
  autolayer(NN_for7, series="NN + Temp + Humidity",PI=FALSE)+
  ylab("Electricity Demand") +
  theme_bw()

# Plot Forecasting Results - Testing Data Only
autoplot(ts_demand_test) +
  autolayer(ARIMA_Fourier_forecast, series="ARIMA + FOURIER",PI=FALSE) +
  autolayer(temphumidity_forecast, series="ARIMA + Fourier + Temp + Humidity",PI=FALSE) +
  autolayer(temponly_forecast, series="ARIMA + Fourier + Temp",PI=FALSE) +
  autolayer(humidity_forecast, series="ARIMA + Fourier + Humidity",PI=FALSE) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE)+
  autolayer(NN_for, series="NN(p=2,P=2)",PI=FALSE)+
  autolayer(NN_for6, series="NN(p=2,P=0)",PI=FALSE)+
  autolayer(NN_for7, series="NN + Temp + Humidity",PI=FALSE)+
  ylab("Electricity Demand") +
  theme_bw()

```
When forecasting on the testing data for 2010, the best model was a neural network with temperature and humidity included. The second-best model in terms of RMSE and MAPE was the ARIMA with exogenous variables. The exogenous variables seemed to really improve the models when forecasting on the testing data, which uses observed values of temperature and humidity. However, these models did not perform as well in the Kaggle competition. This is likely due to the simple methods we used to project temperature and humidity values in 2011. In the Kaggle competition, the neural network models without exogenous variables performed the best. After the models with exogenous variables, the two NN models without exogenous variables had the next lowest RMSE and MAPE when evaluating on the testing data. 
